Effect of Election on Currency Fluctuation

### 2024-04-15 paper/paper.qmd
finished drafting the data section

USD is one of the most popular currency in circulation today. As a foreign buyer of USD product, the final price can vary greatly depending on the exchange rate between one's local currency and USD. Fluctuation of exchange rate can be attributed to multiple different factors. For example the difference in interest rate between two countries, inflation rate, etc. This paper will focus on one particular aspect that contributes to the change in USD exchange rate: the expectation of the populous on the future of the United States, perceived through the results of the US Presidential Election.

@@

@@

## Data on Election Results

### Selection of Data Source

@@

knitr::kable(cleaned_inauguration, col.names = c("President", "Party", "Inauguration Date", "Change Party"), align = c("c", "c", "c", "c"))
```

### Construction of Inauguration Week
$$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  \textbf{Inauguration Day} \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$
```{r}
#| label: tbl-exchange_inaug
#| tbl-cap: Sample of the final dataset used for analysis
#| echo: false
sample_tbl <- exchange_inaug |> filter(date >= "1977-01-12" & date <= "1977-01-18")

@@

## Explaination for small R^2
```{r}
#| label: fig-size_visual
#| fig-cap: Estimate average change in Exchange Rate
#| echo: false
#| warning: false
#| message: false
colors <- c("No Inauguration Announcement" = "yellow", "Inauguration Announcement" = "blue")

@@

# Appendix {-}

>>>>

USD is one of the most popular currency in circulation today. As a foreign buyer of USD product, the final price can vary greatly depending on the exchange rate between one's local currency and USD. Fluctuation of exchange rate can be attributed to multiple different factors. For example the difference in interest rate between two countries, inflation rate, etc. This paper will focus on one particular aspect that contributes to the change in USD exchange rate: the expectation of the populous on the future of the United States, perceived through the results of the US Presidential Election. The election of a new president is often one of the biggest political shift for a country, shifts in stratergies and focus will cause many people to reevaluate their faith for their country.

@@

The DEXCAUS [@Federal_Reserve] dataset includes the daily exchange rate between USD and CAD (base USD = 1) from 1971-01-04 to\
2024-03-29 (date the dataset is downloaded for the purpose of the paper). Raw data contains only the date and the corresponding exchange rate of the day. Exchange rate in this dataset is used as the estimate of this paper.

@@

## Data on Election Results {#sec-election}

In order to measure the effect that the population's faith to their country have on exchange rate, this paper focuses on one particular political event that will quickly change the population's expectation for their country in a short amount of time, the presidential election. Analysis will be focused on the election result announcement and the inauguration period (please see specific definition of the term inauguration period in the subset below @sec-inauguration). Data of past election results and inauguration dates are obtained through the US Census Bureau [@Census] and cross-referenced with Frank LaRose Ohio Secretary of State [@Ohio] for accuracy. The final dataset use the inauguration dates and political party information to determine the effect of having a new president on the USD exchange rate.

@@

select_inauguration <- cleaned_inauguration |> select(-change_party)

knitr::kable(select_inauguration, col.names = c("President", "Party", "Inauguration Date"), align = c("c", "c", "c"))
```

### Construction of Inauguration Week {#sec-inauguration}

Using the inauguration date information present in the dataset, we will construct an announcement period, allowing the news of the result to sink in and give time for the exchange rate to react to the announcement. This period needs to be relatively short to ensure that the change in exchange rate is causal by the election result announcement and **only** the announcement.

The final construction will be referred to as the *inauguration week* for the remainder of this paper. The inauguration week is a +/-3 days week centered on the original inauguration date for each past US president. $$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  \textbf{Inauguration Day} \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$

## Final Dataset
The final dataset combines the date and exchange rate columns from the DEXCAUS [@Federal_Reserve] dataset and uses the inauguration week (constructed through the inauguration dates from the election results dataset, @sec-election). The column *Inauguration Period* has a value of 1 if the date is within the inauguration week and 0 otherwise. A new column *Change Party* is added, it receives a value of 1 if the political party of the new president is different from the previous president's and 0 otherwise.
```{r}
#| label: tbl-exchange_inaug
#| tbl-cap: Sample of the final dataset used for analysis
#| echo: false
sample_tbl <- exchange_inaug |> filter(date >= "1977-01-12" & date <= "1977-01-18")

@@

## Explaination for small R\^2

```{r}
#| label: fig-size_visual
#| fig-cap: Estimate average change in Exchange Rate
#| echo: false
#| warning: false
#| message: false
colors <- c("No Inauguration Announcement" = "yellow", "Inauguration Announcement" = "blue")

@@

# Appendix {.unnumbered}
